---
title: Reverse geocoding from coordinate to ZCTA in R
author: Ran Li
date: '2021-07-23'
slug: reverse-geocoding-from-coordinate-to-zcta-in-r
categories: []
tags: []
subtitle: ''
summary: ''
authors: []
lastmod: '2021-07-23T10:05:05-04:00'
featured: no
image:
  caption: ''
  focal_point: ''
  preview_only: no
projects: []
editor_options: 
  chunk_output_type: console
---

```{r message=FALSE, warning=FALSE, include=FALSE}
library(tidyverse)
library(shiny)
library(sf)
library(rgdal)
library(purrr)
library(knitr)
library(leaflet)
library(glue)
load("sf_zcta_philly.rdata")
df_coord = sf_zcta_philly %>% 
  st_centroid() %>% 
  st_coordinates() %>% 
  as_tibble() %>% 
  mutate(zcta_origin = sf_zcta_philly$zcta ) %>% 
  sample_n(5) %>% 
  mutate(id = row_number()) %>% 
  select(id, longitude=X,  latitude = Y,zcta_origin)  

```

# Intro

[**Goal**]{.ul}: Starting with a list of coordinates (lng and lat) we want to find which ZCTA they belong to.

[**Data required**]{.ul}:

-   Table of coordinates

-    ZCTA boundaries/shape files: You can find the most shape files for US geographic units at the [US Census Bureau website](https://www.census.gov/geographies/mapping-files/time-series/geo/tiger-line-file.html), the specific link for ZCTA files are. [here](https://www2.census.gov/geo/tiger/TIGER2019/ZCTA5/).

[**Libraries require**]{.ul}:

-   tidyverse: for data management with dplr and functional programming tools with purrr.

-   sf: working with [simple features](https://r-spatial.github.io/sf/).

-   rgdal: for reading shape files.

# Step 1: Prepare your data

The code below loads libraries and the data required. Note for this example we providing a minumum example looking at ZCTAs in Philadelphia. We first get some coordinates by taking the centroid of the ZCTA and taking a sample of 5 ZCTA cetrnoids; these will be used for reverse geocoding. The answer is stored in the column 'zcta_origin'. You will probably have some coordinates already in hand... so just use your table here.

```{r eval=FALSE, warning=FALSE}
### Loadl libraries
library(tidyverse)
library(sf)
library(rgdal)

### Read in ZCTA shape files
sf_zcta_philly = readOGR("Census ZCTA Shape Files/tl_2019_us_zcta510.shp") %>%
  st_as_sf()   %>% 
  mutate(zcta = ZCTA5CE10 ) %>% 
  left_join(xwalk_zcta_bchc) %>% 
  filter(city == "Philadelphia")

### Get centroids/coords for us to do reverse geocoding. 
df_coord = sf_zcta_philly %>% 
  st_centroid() %>% 
  st_coordinates() %>% 
  as_tibble() %>% 
  mutate(zcta_origin = sf_zcta_philly$zcta ) %>% 
  sample_n(5) %>% 
  mutate(id = row_number()) %>% 
  select(id, longitude=X,  latitude = Y,zcta_origin)  
```

The table of coordinates we are work with looks like this... just five rows for this example.

```{r echo=FALSE, warning=FALSE}
df_coord%>% 
  knitr::kable()
```

# Step 2: Function to find out which ZCTA they intersect with

Lets first write some code to see which ZCTA a coordinate pair intersects with.

```{r warning=FALSE}
### Lets take the first row and make a  point object
lng_tmp = df_coord %>% slice(1) %>% pull(longitude)
lat_tmp = df_coord %>% slice(1) %>% pull(latitude)
zcta_origin_tmp = df_coord %>% slice(1) %>% pull(zcta_origin)
pt_tmp = st_point(x = c(lng_tmp, lat_tmp), dim = "XYZ")

### see which polygon/row of sf_zcta_philly this point intersects with
intersection_row_tmp = st_intersects(pt_tmp,sf_zcta_philly)[[1]]
sf_zcta_tmp = sf_zcta_philly %>% slice(intersection_row_tmp)
zcta_tmp = sf_zcta_tmp$ZCTA5CE10

### Lets check by plotting the starting coordinate and the ZCTA we matched
leaflet() %>% 
  addTiles() %>% 
  addMarkers(lng = lng_tmp, lat = lat_tmp,  
             label = glue("ZCTA_origin: {zcta_origin_tmp}</br>Coord: ({round(lng_tmp,2)}, {round(lat_tmp,2)})") %>% HTML(),
             labelOptions = labelOptions(noHide = T, direction = "bottom")) %>% 
  addPolygons(data = sf_zcta_tmp, label = paste("ZCTA:",zcta_tmp))

```

Indeed the map shows that we have correctly matched coordinate to ZCTA. Moreover, if you hover over the polygon you can see that the ZCTA matches the ZCTA_origin of the point. Now since we want to iterate this process over many points. Lets turn the above process into a function.

```{r warning=FALSE}
get_zcta_from_coord = function(row_tmp){
  ## Slice row/coord of interest
  df_coord_row = df_coord %>% slice(row_tmp)
  ### Make a  point object
  pt_tmp = st_point(x = c(df_coord_row$longitude, df_coord_row$latitude), dim = "XYZ")
  ### see which polygon/row of sf_zcta_philly this point intersects with
  intersection_row_tmp = st_intersects(pt_tmp,sf_zcta_philly)[[1]]
  sf_zcta_tmp = sf_zcta_philly %>% slice(intersection_row_tmp)
  zcta_tmp = sf_zcta_tmp$ZCTA5CE10
  ### Return table with id and matched zcta
  tibble(id = df_coord_row$id, 
         zcta_matched = zcta_tmp)
}

```

# Step 3: Repeat process for each coordinate

The next step is to iterate or loop over each coordiante we want to find ZCTA's for. We will use a functional programming approach with the map() from purrr as it makes iterating much easier. So we will pass in a vector of numbers 1 to number of rows in df_coord. Then we will run get_zcta_from_coord() for each coord or row in df_coord. With one line of code we can get the final result.

```{r warning=FALSE}
df_results = map_df(1:nrow(df_coord),get_zcta_from_coord)
df_results %>% kable()
```

Lets check the final result with our answers aka zcta_origin

```{r warning=FALSE}
df_results %>%
  left_join(df_coord %>% select(id, zcta_origin)) %>% 
  kable()
```

The ZCTAs that we matched with our function indeed match the original ZCTA. :smile: :taco:
